---
title: "Final Project"
author: "Yiwei Li"
date: "2022-11-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Data Description 

- **Timestamp**
- **Age**
- **Gender**
- **Country**
- **state**: If you live in the United States, which state or territory do you live in?
- **self_employed**: Are you self-employed?
- **family_history**: Do you have a family history of mental illness?
- **treatment**: Have you sought treatment for a mental health condition?
- **work_interfere**: If you have a mental health condition, do you feel that it interferes with your work?
- **no_employees**: How many employees does your company or organization have?
- **remote_work**: Do you work remotely (outside of an office) at least 50% of the time?
- **tech_company**: Is your employer primarily a tech company/organization?
- **benefits**: Does your employer provide mental health benefits?
- **care_options**: Do you know the options for mental health care your employer provides?
- **wellness_program**: Has your employer ever discussed mental health as part of an employee wellness program?
- **seek_help**: Does your employer provide resources to learn more about mental health issues and how to seek help?
- **anonymity**: Is your anonymity protected if you choose to take advantage of mental health or substance abuse treatment resources?
- **leave**: How easy is it for you to take medical leave for a mental health condition?
- **mentalhealthconsequence**: Do you think that discussing a mental health issue with your employer would have negative consequences?
- **physhealthconsequence**: Do you think that discussing a physical health issue with your employer would have negative consequences?
- **coworkers**: Would you be willing to discuss a mental health issue with your coworkers?
- **supervisor**: Would you be willing to discuss a mental health issue with your direct supervisor(s)?
- **mentalhealthinterview**: Would you bring up a mental health issue with a potential employer in an interview?
- **physhealthinterview**: Would you bring up a physical health issue with a potential employer in an interview?
- **mentalvsphysical**: Do you feel that your employer takes mental health as seriously as physical health?
- **obs_consequence**: Have you heard of or observed negative consequences for coworkers with mental health conditions in your workplace?
- **comments**: Any additional notes or comments


***

## Data Preprocessing


```{r}
# read in data
Survey <- read.csv("Data/survey.csv")

time <- Survey['Timestamp']
comments <- Survey['comments']

Survey <- Survey[,2:26]

Survey[sapply(Survey, is.character)] <- lapply(Survey[sapply(Survey, is.character)], as.factor)

Survey <- cbind(time, Survey, comments)

# summary
summary(Survey)
```

We need to convert **Gender** to Male, Female, and Other; make sure **Age** is within the correct range; and remove countries with a **State** that is not US.

```{r}
unique(Survey$Gender)
```

```{r}
Survey_cleaned <- Survey

## convert gender
gender <- Survey_cleaned$Gender
gender_new <- rep("Other", length(gender))

# male
gender_new[grepl("^m.*e$|^M|(Male)|(Man)|(Guy)|(m)|(Mail)", gender)] <- "Male"
# female
gender_new[grepl("[fF]|[wW]|(she)", gender)] <- "Female"


Survey_cleaned$Gender <- as.factor(gender_new)


## check age
age <- Survey_cleaned$Age
for (i in 1:length(age)) {
  if (age[i] < 18 || age[i] >= 90) {
    age[i] <- NA
  }
}
Survey_cleaned$Age <- age


# remove state in non-US countries
Survey_cleaned[Survey_cleaned$Country != "United States" & !is.na(Survey_cleaned$state),]$state <- NA


# summary
summary(Survey_cleaned)
```

In order to build model to analysis the data, we need to convert each categorical data to numerical values. We will use the following scale for the conversion.

- **Gender**: 1 $\rightarrow$ Female, 2 $\rightarrow$ Male, 3 $\rightarrow$ Other
- **self_employed**: No $\rightarrow$ 1, Yes $\rightarrow$ 2
- **family_history**: No $\rightarrow$ 1, Yes $\rightarrow$ 2
- **treatment**: No $\rightarrow$ 1, Yes $\rightarrow$ 2
- **work_interfere**: Never $\rightarrow$ 1, Often $\rightarrow$ 2, Rarely $\rightarrow$ 3, Sometimes $\rightarrow$ 4
- **no_employees**: 1-5 $\rightarrow$ 1, 100-500 $\rightarrow$ 2, 26-100 $\rightarrow$ 3, 500-1000 $\rightarrow$ 4, 6-25 $\rightarrow$ 5, More than 1000 $\rightarrow$ 6
- **remote_work**: No $\rightarrow$ 1, Yes $\rightarrow$ 2
- **tech_company**: No $\rightarrow$ 1, Yes $\rightarrow$ 2
- **benefits**: Don't know $\rightarrow$ 1, No $\rightarrow$ 2, Yes $\rightarrow$ 3
- **care_options**: No $\rightarrow$ 1, Not sure $\rightarrow$ 2, Yes $\rightarrow$ 3
- **wellness_program**: Don't know $\rightarrow$ 1, No $\rightarrow$ 2, Yes $\rightarrow$ 3
- **seek_help**: Don't know $\rightarrow$ 1, No $\rightarrow$ 2, Yes $\rightarrow$ 3
- **anonymity**: Don't know $\rightarrow$ 1, No $\rightarrow$ 2, Yes $\rightarrow$ 3
- **leave**: Don't know $\rightarrow$ 1, Somewaht difficult $\rightarrow$ 2, Somewhat easy $\rightarrow$ 3, Very difficult $\rightarrow$ 4, Very easy $\rightarrow$ 5
- **mentalhealthconsequence**: Maybe $\rightarrow$ 1, No $\rightarrow$ 2, Yes $\rightarrow$ 3
- **physhealthconsequence**: Maybe $\rightarrow$ 1, No $\rightarrow$ 2, Yes $\rightarrow$ 3
- **coworkers**: No $\rightarrow$ 1, Some of them $\rightarrow$ 2, Yes $\rightarrow$ 3
- **supervisor**: No $\rightarrow$ 1, Some of them $\rightarrow$ 2, Yes $\rightarrow$ 3
- **mentalhealthinterview**: Maybe $\rightarrow$ 1, No $\rightarrow$ 2, Yes $\rightarrow$ 3
- **physhealthinterview**: Maybe $\rightarrow$ 1, No $\rightarrow$ 2, Yes $\rightarrow$ 3
- **mentalvsphysical**: Don't know $\rightarrow$ 1, No $\rightarrow$ 2, Yes $\rightarrow$ 3
- **obs_consequence**: No $\rightarrow$ 1, Yes $\rightarrow$ 2



```{r}
Survey_num <- Survey_cleaned
Survey_num[sapply(Survey_num, is.factor)] <- lapply(Survey_num[sapply(Survey_num, is.factor)], as.numeric)

Survey_num$state <- Survey_cleaned$state
Survey_num$Country <- Survey_cleaned$Country
Survey_num$Age <- as.integer(Survey_num$Age)
```


***

## EDA

1. Distribution of Age

```{r}
library(ggplot2)
library(scales)

ggplot(Survey_num, aes(x=Age)) +
  geom_histogram(aes(y = ..density..), color = "white", fill = "#edcb69") +
  geom_density(color = "#385a93", size = 1) +
  labs(x = "Age", y = "Density") +
  scale_x_continuous(limits = c(15,75), breaks = scales::pretty_breaks(n = 5)) + 
  theme(panel.background = element_rect(fill = "#f5f0ec"),
        plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
        panel.grid.minor = element_blank())
```

2. Correlation Matrix

```{r}
library(reshape2)

# use non-na value (971 observations)
corr <- cor(na.omit(Survey_num[, c(2,3,6:26)]))

# reshape to long format
corr <- melt(corr)

length(unique(corr$value))

palette <- colorRampPalette(c("#385a93","#f5f0ec","#385a93"))

ggplot(corr, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white", lwd = 0.5) +
  scale_fill_gradient2(low = "#385a93", mid = "#f5f0ec", high = "#385a93") +
  guides(fill = guide_colorbar(barwidth = 0.5, title = "")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        axis.text.y = element_text(size = 7),
        plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
        legend.background = element_rect(fill = "#f5f0ec")) +
  coord_fixed() +
  labs(x = NULL, y = NULL)


  
```

3. Mental Health Issues within the US

```{r}
#https://www.littlemissdata.com/blog/usmap
#https://stackoverflow.com/questions/60806822/how-do-i-change-state-or-counties-label-sizes-in-r-with-the-function-usmap

library(usmap)
library(dplyr)

Survey_US <- Survey_num[Survey_num$Country == "United States", ]

# with mental health issues -> treatment == 2
stat <- Survey_US %>% 
  group_by(state) %>% 
  summarize(count=sum(treatment==2), percent=sum(treatment==2)/n())
# remove NA
stat <- stat[1:45,]

# plot count
p <- plot_usmap(data = stat, value = "count", labels = T, color = "#f5f0ec", size = 0.6) +
  scale_fill_continuous(low = "white", high = "#385a93", name = "Count") +
  theme(legend.position = "right",
        plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
        legend.background = element_rect(fill = "#f5f0ec"))
# change label size
p$layers[[2]]$aes_params$size <- 3
p

# plot percentage
p <- plot_usmap(data = stat, value = "percent", labels = T, color = "#f5f0ec", size = 0.6) +
  scale_fill_continuous(low = "white", high = "#385a93", name = "Count") +
  theme(legend.position = "right",
        plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
        legend.background = element_rect(fill = "#f5f0ec"))
# change label size
p$layers[[2]]$aes_params$size <- 3
p
```

```{r}
stat
```








***


## Statistical Test

Given the data, we are interested in the following questions:

1. If there is any difference of getting depression or mental health problems between self employed and non-self employed people
2. Does remote work increase or decrease the change of getting mental health problems
3. Is there any association between the number of employees and the chance of developing mental health problems
4. How's the current mental health issues among the tech companies in the US compared with other countries



```{r}
# 1. If there is any difference of getting depression or mental health problem between self employed and non-self employed people
# self employed -> 2, not self employed -> 1
# with mental health issues -> 2, not -> 1
# two proportions test

Q1 <- na.omit(Survey_num[,c("self_employed","treatment")])
Q1_self <- Q1[Q1$self_employed == 2, ]
Q1_not <- Q1[Q1$self_employed == 1, ]


test1 <- prop.test(x = c(sum(Q1_self$treatment == 2), sum(Q1_not$treatment == 2)), n =c(nrow(Q1_self), nrow(Q1_not)))
test1
```


```{r}
#2. Does remote work increase or decrease the change of getting mental health problems
# remote -> 2, not remote -> 1
# with mental health issues -> 2, not -> 1
# two proportions test

Q2 <- na.omit(Survey_num[,c("remote_work","treatment")])
Q2_remote <- Q2[Q2$remote_work == 2, ]
Q2_not <- Q2[Q2$remote_work == 1, ]


prop.test(x = c(sum(Q2_remote$treatment == 2), sum(Q2_not$treatment == 2)), n =c(nrow(Q2_remote), nrow(Q2_not)))
```


```{r}
#3. Is there any association between the number of employees and the chance of developing mental health problems
# 1-5 -> 1, 100-500 -> 2, 26-100 -> 3, 500-1000 -> 4, 6-25 -> 5, >1000 -> 6
# with mental health issues -> 2, not -> 1
# multiple proportions test using marascuilo procedure

Q3 <- na.omit(Survey_num[, c("no_employees", "treatment")])

df <- Q3 %>% 
  group_by(no_employees) %>% 
  summarize(count = sum(treatment == 2), total = n())
df

df2 <- data.frame("G1"=c(162,91), "G2"=c(176,95), "G3"=c(289,150), "G4"=c(60,27), "G5"=c(290,128), "G6"=c(282,146))
rownames(df2) <- c("Y","N")
colnames(df2) <- c("1-5","6-25","26-100","100-500","500-1000",">1000")
df2

marascuilo(df2)
```


```{r}
# http://datalackey.com/2019/04/16/assessing-which-of-a-series-of-pairwise-combinations-significantly-differ-via-the-marascuilo-procedure-in-r/
marascuilo = function(dataFrame,confidence=.95) {

 chiResult = chisq.test (dataFrame, correct=FALSE )
 xSquared = chiResult$statistic

 # Generate all possible pair-wise combinations of groups
 colNames = names(dataFrame)
 combos = combn(colNames , 2)
 numCombos = dim(combos)[2]  # combos is an array of pairs, we want the length


 # Allocate matrix (initially 0 rows) for results
 results = matrix(nrow=0, ncol=5, dimnames=getResultsColumNames() )

 chiSquaredConstant = calcChiSquaredConstant(dataFrame, confidence)
 for (i in 1: numCombos) { 
   newRow = testSignificanceOfAbsDiffVsCriticalRange(
                        dataFrame, combos, i, chiSquaredConstant ) 
    results = rbind(results, newRow)        # append new row to results
 }


 # sort results so that the pair differences that most strikingly exceed 
 # the critical range appear toward the top.
 sortedResults = results[  order( results[,'abs.diff-critical.range'] ) , ]
 return (sortedResults )
}


calcChiSquaredConstant = function(dataFrame,confidence) {
  nRows = dim(dataFrame)[1]  
  nCols = dim(dataFrame)[2]  

  degreesFreedom =  (nRows-1) * (nCols-1) 
  chiSquaredConstant = sqrt( qchisq(confidence,degreesFreedom) )

  return (chiSquaredConstant)
}


getResultsColumNames =  function (numRows) {
   return ( 
        list( 
            c(), 
            c('pair', 'abs.diff', 'critical.range', 'abs.diff-critical.range', 'significant')
        ) 
   )
}

# test significance for ith combination
#
testSignificanceOfAbsDiffVsCriticalRange = function(
                dataFrame, combos, i,  chiSquaredConstant) {

   results = matrix(nrow=1, ncol=5, dimnames=getResultsColumNames() )

   pair1=combos[1,i]
   pair2=combos[2,i]

   # sum column denoted by name 'pair1' into groupTotal1 
   groupTotal1 = sum( dataFrame[ , pair1])  
   groupTotal2 = sum( dataFrame[ , pair2])  # do same thing for pair2... 

   p1 = dataFrame[1, pair1] / groupTotal1 
   p2 = dataFrame[1, pair2] / groupTotal2
   p1Not = (1 - p1)
   p2Not = (1 - p2)

    absDiff = abs( p2  - p1 )

    criticalRange = chiSquaredConstant  * 
                        sqrt(p1*p1Not/groupTotal1 + p2*p2Not/groupTotal2)
    results[1, 'pair'] = paste(pair1,"|",pair2) 
    results[1, 'abs.diff'] = round(absDiff,3)
    results[1, 'critical.range'] = round(criticalRange ,3)
    results[1, 'abs.diff-critical.range'] = round(absDiff - criticalRange ,3)


    if (absDiff > criticalRange) {
        results[1, 'significant'] = 'Y'
    } else {
        results[1, 'significant'] = 'N'
    } 

    return(results)
}
```



```{r}
#4. How's the current mental health issues among the tech companies in the US compared with other countries
# with mental health issues -> 2, not -> 1
# two proportion test

Q4_US <- Survey_num[Survey_num$Country == "United States", c("Country","treatment")]
Q4_Other <- Survey_num[Survey_num$Country != "United States", c("Country","treatment")]

prop.test(x = c(sum(Q4_US$treatment == 2), sum(Q4_Other$treatment == 2)), n =c(nrow(Q4_US), nrow(Q4_Other)))
```








































