---
title: "Final Project"
author: "Yinzhou Wang"
date: "2022-11-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\section{Introduction}
According to the CDC, in the U.S., about half the population will be diagnosed with mental illness at some point in their life. In a given year, 1 in 5 Americans experience a mental disorder, and 1 in 25 Americans experience a severe mental disorder
This data is from U.S. Department of Health and Human Services, containing data of
clients who have used mental health services in facilities that report to individual state
administrative data systems.
```{r}
#summary(df)
library(tidyverse)
library(ggplot2)
```

\section{data cleaning}
```{r}
#select variables of interest
df_1 = df %>% select("AGE", "RACE", "GENDER", "MARSTAT", "EMPLOY", "MH1")
```

Define Variables(-9 means missing data):
AGE: Calculated from the client's date of birth at midpoint of the state's elected reporting period. 14 categories
RACE: Specifies the client's most recent reported race at the end of the reporting period
SEX: Identifies the client's most recent reported sex at the end of the reporting period.
MARSTAT: Identifies the client's marital status.
MH1:Specifies the client's current first mental health diagnosis during the reporting period.

\section{Mental Disorders}
```{r}
# define disorder group
disorder_group = c("Trauma- and stressor-related disorders", "Anxiety disorders", "Attention deficit/hyperactivity disorder (ADHD)", "Conduct disorders", "Delirium, dementia", "Bipolar disorders", "Depressive disorders", "Oppositional defiant disorders", "Pervasive developmental disorders", "Personality disorders", "Schizophrenia or other psychotic disorders", "Alcohol or substance use disorders", "Other disorders/conditions")
# exclude the missing data and first age group
df_1_mh_cleaned = df_1 %>% filter(MH1 != -9)
df_1_mh_cleaned$MH1 = as.factor(df_1_mh_cleaned$MH1)
#plot
ggplot(df_1_mh_cleaned, aes(MH1, fill=MH1)) + geom_bar()+
  labs(title="Mental Disorders Count",
       y="Count", x="Disorders") +
  scale_fill_discrete(labels=disorder_group)+
  theme(panel.background = element_rect(fill = "#f5f0ec"),
        plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
        panel.grid.minor = element_blank(),
        legend.background = element_rect(fill = "#f5f0ec"))
ggsave('disorder_distribution.png',device="png")
```


\section{Age}
```{r}
# define age group
age_group = c("12–14 years", "15–17 years", "18–20 years", "21–24 years", "25–29 years", "30–34 years", "35–39 years", "40–44 years", "45–49 years", "50–54 years", "55–59 years", "60–64 years", "65 years and older")
# summarize by age group
col_age = df_1 %>% select("AGE") %>% group_by(AGE) %>% tally()
# exclude the missing data and first age group
df_1_age_cleaned = df_1 %>% filter(AGE != -9) %>% filter(AGE != 1)
df_1_age_cleaned$AGE = as.factor(df_1_age_cleaned$AGE)
#plot
ggplot(df_1_age_cleaned, aes(AGE, fill=AGE)) + geom_bar()+
  labs(title="Age Distribution",
       y="Count", x="Age Group") +
  scale_fill_discrete(labels=age_group) +
  theme(panel.background = element_rect(fill = "#f5f0ec"),
      plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
      panel.grid.minor = element_blank(),
      legend.background = element_rect(fill = "#f5f0ec"))
ggsave('age_distribution.png',device="png")
```
\section{Race}
```{r}
# exclude the null data
df_1_race_cleaned = df_1 %>% select(RACE)%>% filter(RACE != -9) %>% group_by(RACE) %>% tally()
# define race group
race_group = c("American Indian/Alaska Native", "Asian", "Black or African American", "Native Hawaiian or Other Pacific Islander", "White", "Some other race alone/two or more races")
# plot
ggplot(df_1_race_cleaned, aes(x = "", y = n, fill = factor(RACE))) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  scale_fill_discrete(labels=race_group)+
  labs(fill='Race') +
  theme(panel.background = element_rect(fill = "#f5f0ec"),
    plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
    panel.grid.minor = element_blank(),
    legend.background = element_rect(fill = "#f5f0ec"))
ggsave('Race_distribution.png',device="png")
```
\section{Race Standardized}

```{r}
# summary data from U.S census
standardization_race = pop_2019*c(0.007,0.056,0.122,0.002,0.601,0.028)
df_1_race_cleaned$n = df_1_race_cleaned$n/standardization_race
# plot
ggplot(df_1_race_cleaned, aes(x=RACE,y=n, fill=factor(RACE))) + geom_bar(stat = 'identity')+
  labs(title="Standardized Race",
       y="Percentage", x="Race") +
  scale_fill_discrete(labels=race_group)+
  theme(panel.background = element_rect(fill = "#f5f0ec"),
      plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
      panel.grid.minor = element_blank(),
      legend.background = element_rect(fill = "#f5f0ec"))
ggsave('standardized_race.png',device="png")
```


\section{Gender}
```{r}
# exclude the null data
df_1_gender_cleaned = df_1 %>% select(GENDER) %>% filter(GENDER != -9) %>% group_by(GENDER) %>% tally()
# define Gender group
gender_group = c("Male", "Female")
# plot
ggplot(df_1_gender_cleaned, aes(x = "", y = n, fill = factor(GENDER))) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  scale_fill_discrete(labels=gender_group)+
  labs(fill='Gender')+
  theme(panel.background = element_rect(fill = "#f5f0ec"),
    plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
    panel.grid.minor = element_blank(),
    legend.background = element_rect(fill = "#f5f0ec"))
ggsave('Gender_distribution.png',device="png")
```
\section{Marital Status}
```{r}
# exclude the null data
df_1_marital_cleaned = df_1 %>% select(MARSTAT)%>% filter(MARSTAT != -9) %>% group_by(MARSTAT) %>% tally()
# define Gender group
Marital_group = c("Never married", "Now married", "Separated", "Divorced, widowed")
# plot
ggplot(df_1_marital_cleaned, aes(x = "", y = n, fill = factor(MARSTAT))) +
  geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0)+
  scale_fill_discrete(labels=Marital_group)+
  labs(fill='Marital Status')+
  theme(panel.background = element_rect(fill = "#f5f0ec"),
    plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
    panel.grid.minor = element_blank(),
    legend.background = element_rect(fill = "#f5f0ec"))
ggsave('Marital_distribution.png',device="png")
```

\section{Marital Standardized}

```{r}
# summary data from U.S census
standardization_m = pop_2019*c(0.37,0.493,0.016,0.121)
df_1_marital_cleaned$n = df_1_marital_cleaned$n/standardization_m
# plot
ggplot(df_1_marital_cleaned, aes(x=MARSTAT,y=n, fill=factor(MARSTAT))) + geom_bar(stat = 'identity')+
  labs(title="Standardized Marital Status",
       y="Percentage", x="Status") +
  scale_fill_discrete(labels=Marital_group)+
  theme(panel.background = element_rect(fill = "#f5f0ec"),
      plot.background = element_rect(color = "#f5f0ec", fill = "#f5f0ec"),
      panel.grid.minor = element_blank(),
      legend.background = element_rect(fill = "#f5f0ec"))
ggsave('standardized_marital.png',device="png")
```


\section{limitation}
The data is collected on people who uses health services, can't represent the population's mental health state.

\section{Citation}
Substance Abuse andMentalHealthServicesAdministration,Centerfor Behavioral HealthStatistics
and Quality. Mental Health Client-Level Data 2019. Rockville, MD: Substance Abuse and
Mental Health Services Administration, 2021.